Flattering Advice: Avoiding Disappointment in Advice-Giving

### 2024-12-15 manuscript/manuscript.Rmd
citations needed

# Introduction
Advice plays a critical role in both professional and personal decision-making, 
often serving as a crucial tool for navigating complex challenges and improving 
outcomes [@Gino2007; @Harvey1997; @Soll2009]. While prior research has 
extensively examined the *recipients* of advice (e.g., how they evaluate or act 
on recommendations) far less attention has been devoted to understanding the 
motivations and behaviors of the *givers* of advice. In this paper, we shift the 
focus to advisers, proposing that their recommendations are shaped not only by 
accuracy considerations but also by interpersonal concerns. Specifically, we 
propose that advisers factor in advisees' subjective expectations, providing 
flattering advice that aligns with these expectations, even when doing so leads 
to objectively worse recommendations. Notably, this tendency persists in our 
experimental settings, where advisees' expectations are clearly uninformative 
and advisers are financially incentivized for accuracy.

Advice is often viewed as a mechanism to improve decision satisfaction or 
prediction accuracy, assuming both parties prioritize solving the problem at 
hand. However, prior research highlights that interpersonal dynamics can 
complicate these interactions. Advice seekers may avoid asking for help due to 
fear of rejection [@Benabou2022], and people generally have been shown to avoid
information that could lead to disappointment [@Gill2012; @Golman2017]. 
They are motivated to avoid unpleasant truths in an ethical dilemma [@Shalvi2019] 
and dislike those who deliver unwelcome feedback even when they had no role in
the outcome [@John2019]. These tendencies align with models of belief utility 
[@Loewenstein2018], which suggest that individuals experience 
hedonic costs when receiving bad news. A natural reference point for evaluating 
the valence of advice is the advisee's expectation in the absence of advice
[@Koszegi2006]. Consequently, advisers who are aware of expectations may 
strategically adapt their advice to avoid delivering bad news, even at the 
expense of accuracy.

>>>>

# Introduction
Advice plays a critical role in both professional and personal decision-making, 
often serving as a crucial tool for navigating complex challenges and improving 
outcomes [@Gino2007; @Harvey1997; @Soll2009]. While prior research has 
extensively examined the *recipients* of advice (e.g., how they evaluate or act 
on recommendations) far less attention has been devoted to understanding the 
motivations and behaviors of the *givers* of advice (CITES). In this paper, we shift the 
focus to advisers, proposing that their recommendations are shaped not only by 
accuracy considerations but also by interpersonal concerns. Specifically, we 
propose that advisers factor in advisees' subjective expectations, providing 
flattering advice that aligns with these expectations, even when doing so leads 
to objectively worse recommendations. Notably, this tendency persists in our 
experimental settings, where advisees' expectations are clearly uninformative 
and advisers are financially incentivized for accuracy.

Advice is often viewed as a mechanism to improve decision satisfaction or 
prediction accuracy, assuming both parties prioritize solving the problem at 
hand. However, prior research highlights that interpersonal dynamics can 
complicate these interactions (CITE). Advice seekers may avoid asking for help due to 
fear of rejection [@Benabou2022], and people generally have been shown to avoid
information that could lead to disappointment [@Gill2012; @Golman2017]. 
They are motivated to avoid unpleasant truths in an ethical dilemma [@Shalvi2019] 
and dislike those who deliver unwelcome feedback even when they had no role in
the outcome [@John2019]. These tendencies align with models of belief utility 
[@Loewenstein2018], which suggest that individuals experience 
hedonic costs when receiving bad news. A natural reference point for evaluating 
the valence of advice is the advisee's expectation in the absence of advice
[@Koszegi2006]. Consequently, advisers who are aware of expectations may 
strategically adapt their advice to avoid delivering bad news, even at the 
expense of accuracy.
